Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-03-29 PandaOverview.tex
Reduction III

The design of PanDA WMS started in 2005 and its implementation went into
production for the LHC Run 1 on 2009. PanDA was then extended with new
subsystems to be deployed on Run 2, on 2015. In this section, we summarize the
design, architecture, and execution process of the current state of the
practice. We focus on those subsystems and components that are most relevant to
the the integration of PanDA with supercomputers, the main topic of this paper.

@@

@@

PanDA's application model assumes single tasks, workloads and workflows. Tasks
represent a set of operations performed on a set of events, stored in one or
more input files. Tasks are decomposed into jobs, where each job represents the
task's set of operations and a partition of the task's events. Since 2005, a
certain amount of parallelism has been progressively introduced for job
execution~\cite{crooks2012multi} but, so far, no MPI jobs have been considered
for production. Jobs are supposed to be relatively self-contained, capable of
setting up their own execution environment or having a minimal set of common
dependences.

@@

PanDA's user \jhanote{usage} model is based on multitenancy \jhanote{of what?}
and the support of at least two types of HEP users: individual researchers and
groups executing so called `production' workflows. Users are free to submit
tasks and workflows to the PanDA WMS at any point in time, directly or via
dedicated application frameworks. Consistently, PanDA's security model is
based on separation between authentication, authorization and accounting for
both single users and group of users. Both authentication and authorization
are based on digital certificates and the X.509 standard and on the virtual
organization (VO) abstraction.

@@

In PanDA's data model, each event represents a collision captured in the
detector and one or more events can be packaged into files or data source with
other container abstractions. As with jobs, data have both attributes and
states, and some of the attributes are shared between events and jobs.
% so to create associations.
Raw, reconstruction, and simulation data are assumed to be distributed across
multiple storage facilities and managed by the ATLAS Distributed Data Management
(DDM)~\cite{garonne2012atlas}. Datasets required by each job are assumed to be
replicated when necessary over the network, both for input and output data.

@@

% -----------------------------------------------------------------------------
\subsection{Architecture}
\label{ssec:panda_arch}

The implementation of PanDA consists of several interconnected subsystems, most
of them built from off-the-shelf and Open Source components. Sbub-systems
communicate via dedicated API or HTTP messaging and each sub-system is
implemented by one or more modules. Databases are used to store eventful
entities like tasks, jobs and events, and to store information about sites,
resources, logs, and accounting.

The execution of specific tasks of ATLAS workflows on supercomputers involves
four main subsystems of the PanDA architecture: PanDA Server, AutoPyFactory,
PanDA Pilot, and JEDI. Figure~\ref{fig:architecture} shows a diagrammatic
representation of these sub-systems' architecture. For a full list of subsystems
see Ref.~\cite{panda-wiki_url}.

% In the following, we briefly describe the overall architecture of PanDA and
% its components. The description provides all the details required to
% understand how HPC resources have been integrated within PanDA
% (\ref{sec:integration}).

% For a more detailed description of PanDA architecture see~\cite{wiki, papers}.

@@

% Originally PanDA was designed for grid infrastructure... In this
% paper is focussed on the HPC use case; there are differences from
% the grid use case. This paper discusses how PanDA has been adapted
% to execute a workload % on HPC resources. The schematic overview is
% presented for workload X on Titan......

\paragraph{\textbf{PanDA Server}} this subsystem is the central hub service of
PanDA and it operates as a web service~\cite{maeno2011overview}. It runs on
Apache web server, interacting with the back-end database running on separate
machines. PanDA Server implements the Apache worker model, with many independent
processes handling client requests in parallel. Since all state is maintained in
the central database, the PanDA Server instance itself is stateless. Currently,
the production version of PanDA utilizes an Oracle database as backend, but
PanDA Server can also work with the MySQL family of databases.

% . It provides a task queue
% managing all job information centrally. The PanDA server receives jobs through
% the client interface into the task queue, upon which a brokerage module
% operates to prioritize and assign work on the basis of job type, priority,
% input data and its locality, and available CPU resources. The PanDA server

% and communicating via Python APIs

PanDA Server consists of four main components implemented in Python: Task
Buffer, Brokerage, Job Dispatcher, and Data Service. Task Buffer manages PanDA's
global queue by: (i) storing jobs into the global queue; (ii) providing
information about and updating the state of each stored job. The Brokerage
component: (i) matches job attributes with site and pilot attributes; (ii)
manages the dispatch of input data to processing sites; and (iii) derives
PanDA's data pre-placement requirement. The Job Dispatcher: (i) receives
requests from active pilots for new jobs to process; (ii) retrieves from Task
Buffer the jobs with highest priority and that can be executed on the site of
the requesting pilot; (iii) creates a wrapper script for the selected job,
tailored to the site of the requesting pilot; and (iv) monitors the state of
the job execution on the pilot, updating the Task Buffer component. Finally, the
Data Service component: (i) retrieves information from Task Buffer and Brokerage
about the data requirements of each job and the site to which each job has been
assigned; (ii) interacts with the ATLAS DDM to make input files available to the
jobs and manage their output data.

% The job/pilot matching algorithm assigns a weight to each candidate pilot and
% after applying a set of policies, the best ten candidates are taken.

% including  recovering and cleaning up previously failed jobs on that site and

\paragraph{\textbf{PanDA Pilot}} this subsystem implements the execution
environment for PanDA's jobs\cite{nilsson2011atlas}. Implemented in Python, a
PanDA Pilot is launched by a wrapper script submitted to a site. Wrapper scripts
are tailored to the middleware exposed by each site so to maintain PanDA Pilot
site-independent. PanDA Pilot: (i) prepares the execution environment for
PanDA's jobs, collecting information about the compute, data, and network
capabilities of the site; (ii) requests a job for execution to the Job
Dispatcher component of PanDA Server; (iii) initiates and monitors the execution
of pulled jobs, including setting up the job, transferring its input when
required, executing the payload, transferring output, and checking log files and
local space; and (iv) cleans up when the payload has finished to execute.

@@

Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
HTTP from a central Subversion repository. The repository works with an Apache
web server, configured with a memory-based web proxy (Squid). The purpose of the
cache is to reduce the request load on the back-end Subversion server. That
improves performance since only occasional queries trigger a full lookup on the
back-end Subversion system, and most external queries are pulled from memory on
the front-end web server. The benefits of this system include a high-performance
code download service, combined with code updates still being immediately
available as soon as they are committed to source code control.

\paragraph{\textbf{AutoPyFactory}} this subsystem implements a factory to submit
pilots locally and to both grid and cloud
sites~\cite{caballero2012autopyfactory}. Implemented to be executed as a single
demonized process, AutoPyFactory concurrently spawns multiple job submission
processes called APFQueue. Each APFQueue communicates with PanDA Server and the
Brokerage module to collect information about the state of the jobs it manages.
On the basis of this information, an APFQueue submits new pilots to one site.
AutoPyFactory implements functionalities to evaluate the number of pilots
already instantiated on each site, and to apply a consistent and adequate
pressure on their batch queues. Submission is performed via the Condor-G batch
system~\cite{frey2002condor}, enabling submission to multiple type of sites.
AutoPyFactory exposes a dedicated monitoring service, feeding information to the
PanDA Monitoring subsystem. A proxy manager enables certificate-based
authentication and authorization for each site.

@@

\paragraph{\textbf{JEDI}} this subsystem exposes a task abstraction for workload
and workflows specification and interfaces with PanDA Server to manage their
execution on remote sites~\cite{borodin2015scaling}. Implemented as a standalone
service, JEDI partitions tasks to jobs, and binds these jobs to sites via
dedicated queues. JEDI collects information about sites' resource capability,
availability, and performance. Based on this information, JEDI associates sites
with similar capabilities to a queue. Each job is assigned to a queue by
matching the job's requirements to the capability, availability and performance
of the sites of that queue. In this way, JEDI replaces the Brokerage module of
the PanDA Server.
% This module still handles communication between the Task
% Buffer and the Data Service modules.

% -----------------------------------------------------------------------------
\subsection{Execution}
\label{ssec:panda_exec}

@@

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
required users or applications to perform a static conversion between tasks and
jobs: tasks were described as a set of jobs with a fixed number of events and
then submitted to the PanDA Server (Fig.~\ref{fig:architecture}:A1).

>>>>

PanDA is a Workload Management System (WMS) designed to support the execution of
distributed workloads and workflows via pilots. WMS is middleware for
discovering and selecting resources, submitting tasks of workloads and
workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
abstraction that enables multi-level scheduling by decoupling resource
acquisition from tasks scheduling~\cite{turilli2015comprehensive}. Once
implemented, a pilot is scheduled on a site (e.g., by submitting a job or
requesting a virtual machine) and, once active, tasks are scheduled to the
pilot, not to the site's scheduler.

Pilot-enabled WMS are particularly relevant for LHC experiments, where millions
of tasks are executed across multiple sites every month, analyzing and producing
petabytes of data. The design of PanDA WMS started in 2005 and its
implementation went into production for the LHC Run 1 on 2009. PanDA was then
extended with new subsystems to be deployed on Run 2, on 2015. We summarize
PanDA's design, architecture, and execution process, focussing on those
subsystems and components that are most relevant to the the integration of PanDA
with supercomputers.

@@

PanDA's application model assumes tasks, workloads and workflows. Tasks
represent a set of operations performed on a set of events that are stored in
one or more input files. Tasks are decomposed into jobs, where each job
represents the task's set of operations and a partition of the task's events.
Since 2005, a certain amount of parallelism has been progressively introduced
for job execution~\cite{crooks2012multi} but, so far, no MPI jobs have been
considered for production. Jobs are supposed to be relatively self-contained,
capable of setting up their own execution environment or having a minimal set of
common dependences.

PanDA's usage model is based on multitenancy of resources and the support of at
least two types of HEP users: individual researchers and groups executing so
called `production' workflows. Users are free to submit tasks and workflows to
the PanDA WMS at any point in time, directly or via dedicated application
frameworks. Consistently, PanDA's security model is based on separation between
authentication, authorization and accounting for both single users and group of
users. Both authentication and authorization are based on digital certificates
and the X.509 standard and on the virtual organization (VO) abstraction.

In PanDA's data model, each event refers to the recorded or simulated
measurement of a physical event. One or more events can be packaged into files
or other data containers. As with jobs, data have both attributes and states,
and some of the attributes are shared between events and jobs. Raw,
reconstruction, and simulation data are assumed to be distributed across
multiple storage facilities and managed by the ATLAS Distributed Data Management
(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
assumed to be replicated over the network, both for input and output data.

@@

@@

@@

@@

% -----------------------------------------------------------------------------
\subsection{Implementation and Execution}
\label{ssec:panda_arch}

The implementation of PanDA consists of several interconnected subsystems, most
of them built from off-the-shelf and Open Source components. Subsystems
communicate via dedicated API or HTTP messaging, and each subsystem is
implemented by one or more modules. Databases are used to store eventful
entities like tasks, jobs and events, and to store information about sites,
resources, logs, and accounting.

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}) but, at the
moment, they are not relevant to understand how PanDA has been ported to
supercomputers. For a full list of subsystems see Ref.~\cite{panda-wiki_url}.
Figure~\ref{fig:architecture} shows a diagrammatic representation of PanDA main
subsystems, highlighting the execution process of tasks while omitting
monitoring details to improve readability.

@@

% -----------------------------------------------------------------------------
% PANDA SERVER
% -----------------------------------------------------------------------------
% \textbf{PanDA Server} is the central hub service of PanDA WMS and it operates
% as a web service~\cite{maeno2011overview}. It consists of four main components
% implemented in Python: Task Buffer, Brokerage, Job Dispatcher, and Data
% Service. Together, these components implement a global queue where to store
% jobs, share information about jobs and their dataset properties with the other
% subsystems, and update their states at execution time.
% -----------------------------------------------------------------------------

% It runs on Apache web server, interacting with the back-end database running
% on separate machines. PanDA Server implements the Apache worker model, with
% many independent processes handling client requests in parallel. Since all
% state is maintained in the central database, the PanDA Server instance itself
% is stateless. Currently, the production version of PanDA utilizes an Oracle
% database as backend, but PanDA Server can also work with the MySQL family of
% databases.

% . It provides a task queue
% managing all job information centrally. The PanDA server receives jobs through
% the client interface into the task queue, upon which a brokerage module
% operates to prioritize and assign work on the basis of job type, priority,
% input data and its locality, and available CPU resources. The PanDA server
%
% and communicating via Python APIs

% Task Buffer manages PanDA's global queue by: (i) storing jobs into the global
% queue; (ii) providing information about and updating the state of each stored
% job. The Brokerage component: (i) matches job attributes with site and pilot
% attributes; (ii) manages the dispatch of input data to processing sites; and
% (iii) derives PanDA's data pre-placement requirement. The Job Dispatcher: (i)
% receives requests from active pilots for new jobs to process; (ii) retrieves
% from Task Buffer the jobs with highest priority and that can be executed on
% the site of the requesting pilot; (iii) creates a wrapper script for the
% selected job, tailored to the site of the requesting pilot; and (iv) monitors
% the state of the job execution on the pilot, updating the Task Buffer
% component. Finally, the Data Service component: (i) retrieves information from
% Task Buffer and Brokerage about the data requirements of each job and the site
% to which each job has been assigned; (ii) interacts with the ATLAS DDM to make
% input files available to the jobs and manage their output data.

% -----------------------------------------------------------------------------
% PANDA PILOT
% -----------------------------------------------------------------------------
% \textbf{PanDA Pilot} implements the execution environment for PanDA's
% jobs\cite{nilsson2011atlas}. Implemented in Python, a PanDA Pilot is launched
% by a wrapper script submitted to a site. Wrapper scripts are tailored to the
% middleware exposed by each site so to maintain PanDA Pilot site-independent.
% PanDA Pilot pulls a job for execution, sets up the execution environment
% including the required input files, spawns the jobs's executable, monitors its
% execution, and supports the output's staging.
% -----------------------------------------------------------------------------

% t: (i) prepares the execution environment for PanDA's jobs, collecting
% information about the compute, data, and network capabilities of the site;
% (ii) requests a job for execution to the Job Dispatcher component of PanDA
% Server; (iii) initiates and monitors the execution of pulled jobs, including
% setting up the job, transferring its input when required, executing the
% payload, transferring output, and checking log files and local space; and (iv)
% cleans up when the payload has finished to execute.

@@

% Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
% HTTP from a central Subversion repository. The repository works with an Apache
% web server, configured with a memory-based web proxy (Squid). The purpose of
% the cache is to reduce the request load on the back-end Subversion server.
% That improves performance since only occasional queries trigger a full lookup
% on the back-end Subversion system, and most external queries are pulled from
% memory on the front-end web server. The benefits of this system include a
% high-performance code download service, combined with code updates still being
% immediately available as soon as they are committed to source code control.

% -----------------------------------------------------------------------------
% AUTOPYFACTORY
% -----------------------------------------------------------------------------
% \textbf{AutoPyFactory} implements a factory to submit pilots locally and to
% both grid and cloud sites~\cite{caballero2012autopyfactory}. Implemented to be
% executed as a single demonized process, AutoPyFactory concurrently spawns
% multiple processes called APFQueue that, on the base of information about jobs
% queued on the PanDA server and on the state and capabilities of resources,
% submit pilots to multiple sites with Condor-G~\cite{frey2002condor}.
% -----------------------------------------------------------------------------

% Each APFQueue communicates with PanDA Server and the Brokerage module to
% collect information about the state of the jobs it manages. On the basis of
% this information, an APFQueue submits new pilots to one site. AutoPyFactory
% implements functionalities to evaluate the number of pilots already
% instantiated on each site, and to apply a consistent and adequate pressure on
% their batch queues. Submission is performed via the Condor-G batch
% system~\cite{frey2002condor}, enabling submission to multiple type of sites.
% AutoPyFactory exposes a dedicated monitoring service, feeding information to
% the PanDA Monitoring subsystem. A proxy manager enables certificate-based
% authentication and authorization for each site.

@@

% -----------------------------------------------------------------------------
% JEDI
% -----------------------------------------------------------------------------
% \textbf{JEDI} exposes a task abstraction for workload and workflows
% specification and interfaces with PanDA Server to manage their execution on
% remote sites~\cite{borodin2015scaling}. Implemented as a standalone service,
% JEDI partitions tasks to jobs, and binds these jobs to sites via dedicated
% queues. JEDI collects information about sites' resource capability,
% availability, and performance. Based on this information, JEDI associates
% sites with similar capabilities to a queue. JEDI replaces and extends the
% Brokerage module of the PanDA Server.
% -----------------------------------------------------------------------------

% Each job is assigned to a queue by matching the job's requirements to the
% capability, availability and performance of the sites of that queue. In this
% way,

% This module still handles communication between the Task
% Buffer and the Data Service modules.

% -----------------------------------------------------------------------------
% \subsection{Execution}
% \label{ssec:panda_exec}

@@

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
required users or applications to perform a static conversion between tasks and
jobs: tasks were described as a set of jobs with a fixed number of events and
then submitted to the PanDA Server.
